Gradual Soundness: Lessons from Static Python

### 2021-12-14 pj-7.1/paper.tex
pj: abstract

  \paragraph{Context}
  % What is the broad context of the work? What is
  % the importance of the general research area?
  FILL

  \paragraph{Inquiry}
  % What problem or question does the paper
  % address? How has this problem or question been
  % addressed by others (if at all)?
  FILL

  \paragraph{Approach}
  % What was done that unveiled new knowledge?
  FILL

  \paragraph{Knowledge}
  % What new facts were uncovered? If the
  % research was not results oriented, what new
  % capabilities are enabled by the work?
  FILL

  \paragraph{Grounding}
  % What argument, feasibility proof, artifacts,
  % or results and evaluation support this work?
  FILL

  \paragraph{Importance}
  % Why does this work matter?
  FILL

@@

%% https://docs.python.org/3/reference/datamodel.html

>>>>

  \paragraph{Context}
  % What is the broad context of the work? What is
  % the importance of the general research area?
  Gradually-typed languages allow typed and untyped code to interoperate,
  but typically come with some kind of drawback.
  In some languages, the types are unreliable;
  in others, communication across type boundaries can be extremely expensive;
  and still others allow only limited forms of interoperability.
  The research community is actively seeking a sound, fast, and expressive
  approach to gradual typing.

  %% TODO acknowlegde the success of new language / tracing JIT approaches

  \paragraph{Inquiry}
  % What problem or question does the paper
  % address? How has this problem or question been
  % addressed by others (if at all)?
  This paper describes \SP{}, a language developed by engineers at Instagram
  that has proven itself sound, fast, and reasonably expressive in production.
  \SP{}'s approach to gradual types is essentially a programmer-tunable combination of
  the \emph{concrete}\/ and \emph{transient}\/ approaches from the literature.
  Concrete types provide soundness in an efficient but difficult-to-use way.
  Transient types are sound in a shallow sense and easier to use; they help
  to bridge the gap between untyped code and typed concrete code.

  \paragraph{Approach}
  % What was done that unveiled new knowledge?
  In collaboration with the \SP{} team, we have evaluated the language
  as it exists today and developed a model to capture the essence of its
  approach to gradual types.
  We drew upon personal communications, bug reports, and the \SP{}
  regression test suite.

  \paragraph{Knowledge}
  % What new facts were uncovered? If the
  % research was not results oriented, what new
  % capabilities are enabled by the work?
  Our main finding is that the \emph{gradual soundness}\/ that
  arises from a mix of concrete and transient types is an effective
  way to lower the maintenance cost of the concrete approach.
  On a more technical level, this paper contributes two artifacts:
  a model of \SP{} that passes all relevant tests from the \SP{} codebase
  and a performance evaluation of \SP{}.

  \paragraph{Grounding}
  % What argument, feasibility proof, artifacts,
  % or results and evaluation support this work?
  Our model of \SP{} is implemented in PLT Redex and tested against
  the \SP{} regression suite.
  This paper includes a small core of the model to convey the main ideas
  of the \SP{} approach and its soundness.
  Our performance claims are supported by data from both
  a public suite a microbenchmarks and the proprietary Instagram codebase.

  \paragraph{Importance}
  % Why does this work matter?
  \SP{} is the first language that improves performance across the board through
  the addition of sound types to an untyped language.
  %% Prior work on Nom and Dart 2 showed that concrete types are promising.
  Other language designers may wish to replicate its approach,
  especially those who maintain unsound type checkers for untyped languages,
  e.g., JavaScript.

@@

%% https://docs.python.org/3/reference/datamodel.html
%% TODO check Dino email, confirm with C&D that layout conflict is unrelated to vtables
